---
title: "Hands-on Exercise 1"
author: "EEEcy"
---

# Getting Started

## Install and launch R Packages

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## import the data

```{r}
mpsz <- st_read(dsn = "D:/EEEcy/ISSS626-GAA/Hands-on_Ex/Hands-on_Ex01/data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

Importing polyline feature data in shapefile form

```{R}
mpsz <- st_read(dsn = "D:/EEEcy/ISSS626-GAA/Hands-on_Ex/Hands-on_Ex01/data/geospatial", 
                layer = "CyclingPathGazette")
```

Importing GIS data in kml format

```{R}
preschool = st_read("D:/EEEcy/ISSS626-GAA/Hands-on_Ex/Hands-on_Ex01/data/geospatial/PreSchoolsLocation.kml")
```

Working with st_geometry()

```{R}
st_geometry(mpsz)
```

Working with st_geometry()

```{R}
glimpse(mpsz)
```

Working with head()

```{R}
head(mpsz, n=5)  
```

Plotting the Geospatial Data

```{R}
plot(mpsz) 
```

Working with Projection

```{R}
st_crs(mpsz)  
```

```{R}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

```{R}
st_crs(mpsz3414)
```

```{R}
preschool3414 <- st_transform(preschool, 
                              crs = 3414) 
```

Importing the aspatial data

```{R}
listings <- read_csv("Hands-on_Ex01/data/aspatial/listings.csv")
```

```{R}
list(listings) 
```

Creating a simple feature data frame from an aspatial data frame

```{R}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{R}
glimpse(listings_sf)
```


The scenario:

A pre-school service group want to find out the numbers of pre-schools in each Planning Subzone.

```{R}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
top_n(mpsz3414, 1, `PreSch Count`)
```

```{R}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

```{R}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```
